### 2023-01-20 code/may.Rmd

---
  output: github_document
---

  ```{r}
knitr::opts_chunk$set(fig.width = 7)
```

```{r message=FALSE}
#library(MDPtoolbox)
library(tidyverse) # for plotting
```

```{r}
states <- seq(0,2, length=100)
actions <- seq(0,1, length.out = 10)
observations <- states
sigma_g <- 0.2
sigma_m <- 0.0

Q <- 2
V <- 15
C <- 10
reward_fn <- function(x,h) h
discount <- 0.999

r <- 0.8
K <- 1.53
q <- 2
b <- .2
# a <- 0.27 # only one (high) stable state, but creates nonlinear policy
a <-   0.28 # .27 #0.29

f <- function(x, h){ # May
  # h is conservation effort

  x + x * r * (1 - x / K) -  a * x ^ q / (x ^ q + b ^ q) - h * x
}
```

```{r}
# B_MSY - biomass max growth rate
X <- seq(0,1.2, length.out=200)
V <- map_dbl(X, \(x) f(x,0) - x)

tibble(X, V) |> ggplot(aes(X, V)) + geom_point()

i <- which.max(V)
B_MSY <- X[i]

F_MSY <- (f(X[i],0) - X[i]) / X[i]

```

```{r}

df <- map_dfr(seq(0, 1, length.out = 31),
function(init) {
  df <- tibble(init, time = 1:50, x = init)
  for(t in 1:max(df$time)) {
    df[t+1,] = tibble(init, time = t+1, x = f(df$x[t], F_MSY)) 
  }
  df
})
df2 <- df |> mutate(h = F_MSY * x)
df2 |> ggplot(aes(time,x, col=init, group=init)) +
  geom_line() + ylim(0,1) + 
  geom_hline(aes(yintercept=X[i]))
```
```{r}

  df <- map_dfr(seq(0, 1, length.out = 31),
  function(init) {
    df <- tibble(init, time = 1:50, x = init)
    for(t in 1:max(df$time)) {
      effort = pmax( (df$x[t] - B_MSY) /  df$x[t], 0)
      df[t+1,] = tibble(init, time = t+1, x = f(df$x[t], effort)) 
    }
    df
  })
  df |> ggplot(aes(time,x, col=init, group=init)) +
    geom_line() + ylim(0,1)
```

```{r}
df <- 
bind_rows(
  tibble(policy="const mortality", x = seq(0,1.2, length.out=100), delta = f(x,F_MSY) - x),
  tibble(policy="const escapement", x = seq(0,1.2, length.out=100), delta = f(x,0) - x),
  )
df %>% 
  ggplot(aes(x, delta, col=policy, group=policy)) + 
  geom_line() + geom_hline(aes(yintercept=0))
```